---
title: "Location based competition in Edinburgh restaurants"
author: "Francisco Marco-Serrano"
date: "20 November 2015"
output: html_document
---

#Abstract

Location models refer to competition models in economics that link consumer preference for a brand and its location.

The original models are Hotelling's linear city or Salop's circle model, in which the relationship between product characteristics and geographic location determine the competitive landscape; brands compete in terms of location, rather than in price or other characteristics.

Using the Yelp dataset, that includes information on 366K users (consumers), 61K businesses (brands), and 481K business attributes (characteristics than include geographic location), I have tested the empirical application of these economic location models and estimated competition levels in a city.

---

#Introduction

From an economics modelling point of view, competition can be addressed from a perspective of product differentiation. This differentiation can be considered in a multivariate (or multiattribute) facet.

Sometimes, the attributes are quite similar and the only differentiation is quality. However, even this feature can be opaque to the consumer; and, here is when services like Yelp's improve the consumption process, by indicating the quality of a product or service. In our case, we intend to analyse the competition model operating in the Edinburgh restaurants, using the Yelp Academic Dataset - Round 6.

A seminal model to analyse this type of competition is that of Hotelling (1929), expanded later by Salop (1979). In these models, there is what Hotelling named as the 'principle of minimum differentiation', which is resolved by businesses differentiating themselves in one or few attributes, being the one explored by Hotelling and Salop the location of a business.

The aim of this project is to find out if either (a) **location** is an attribute that restaurants in Edinburgh are using to differentiate from competition, (b) **quality** is the differentiation attribute, or (c) there is any other feature used for differentiating from competition.


*Hotelling, H. (1929). "Stability in Competition". Economic Journal, 96(153), 41-57.*
*Salop, S. (1979). "Monopolistic Competition with Outside Goods". Bell Journal of Economics, 10(1), 141-156.*

---

#Methods and Data

The full dataset is made up of 1.6M reviews and 500K tips by 366K users for 61K businesses, 481K business attributes (i.e. hours, parking availability, ambience),
a social network of 366K users, for a total of 2.9M social edges, and the aggregated check-ins over time for each of the 61K businesses.

The dataset can be obtained from: http://www.yelp.com/dataset_challenge

In this project only the Edinburgh businesses categorised as restaurant have been used. As well, the information relating to the social network and check-ins have not been used, leaving it for further research.


The data finally used is:

* Business - 945 observations of 98 variables, including location, opening/closing times, and business attributes. Businesses are ranked by the users on a 1 to 5 star-system. The descriptive statistics for Edinburgh restaurants are:

```{r, echo=FALSE, message=FALSE}

library(jsonlite)

biz <- readRDS("biz.rds")
biz <- biz[which(biz$city == "Edinburgh"), ]
biz <- biz[which(biz$open==TRUE), ]
biz <- biz[which(grepl("Restaurants", biz$categories)), -c(2, 4, 6, 8, 9, 11, 15)]
biz <- flatten(biz, recursive = TRUE)

summary(biz$stars)

```


* Users - 366715 observations of 19 variables, including number of fans, account age, number of votes and compliments received
* Reviews - 10506 observations of variables, including date of the review, votes received, and stars

The three datasets have been merged in a single dataset -master (review-business-user) comprising 10506 observations of 122 variables.

The analysis of the data has been done using random forests (Breiman, 2001). This methodology combines bagging with the random selection of variables to obtain a collection of decision trees with controlled variance. It can be used for classification or regression, and both in supervised and unsupervised mode.


*Breiman, L. (2001). "Random Forests". Machine Learning, 45, 5-32.*

---

#Results

The spatial distribution of the restaurants can be seen here:

```{r, echo=FALSE, message=FALSE}

library(ggmap)

map <- get_map(location = 'Edimburgh', zoom = 11, maptype = "roadmap")
mapPoints <- ggmap(map) +
          geom_point(aes(x = biz$longitude, y = biz$latitude),
                         data = biz, size = 1, alpha = 0.5)
mapPoints
```

And the distribution (in %) of businesses based on their location coordinates:

```{r, echo=FALSE, message=FALSE}

round(table(cut(biz$latitude, quantile(biz$latitude)),
      cut(biz$longitude, quantile(biz$longitude)))/945*100,2)

```


Running an unsupervised random forest on the businesses dataset, to find out the Top 10 variables in importance that classify the businesses:

```{r, echo=FALSE, message=FALSE}
library(randomForest)

#Flatten the data frame to avoid nested structure,
#which comes from the former json nature of the data
biz <- flatten(biz, recursive = TRUE)

#NA in Attributes is considered to be FALSE
#See http://stackoverflow.com/questions/2991514/r-preventing-unlist-to-drop-null-values
#to solve NULL creating problems when flattening list variables
biz$`attributes.Accepts Credit Cards`[sapply(biz$`attributes.Accepts Credit Cards`, is.null)] <- NA
biz$`attributes.Accepts Credit Cards` <- unlist(biz$`attributes.Accepts Credit Cards`,
                                                recursive = T, use.names = T)

attrib <- grepl("attributes.", names(biz))
biz[attrib] <- replace(biz[attrib], is.na(biz[attrib]), FALSE)
biz[attrib] <- lapply(biz[attrib], as.factor)


#NA in Hours is considered to be 00:00
hours <- grepl("hours.", names(biz))
biz[hours] <- replace(biz[hours], is.na(biz[hours]), "00:00")
biz[hours] <- lapply(biz[hours], as.factor)

#Run the random forest
features <- colnames(biz)[-c(1, 2)]
ntrees=3000

set.seed(123)
fit.urf <- randomForest(x=biz[,features], ntree=ntrees, importance=TRUE,mtry=3)

#Top variable importance
varImpPlot(fit.urf,type=2)

```


---

#Discussion

Analysing Yelp Academic Dataset we were trying to test the approach that Edinburgh restaurants were following to differentiate themselves from their near competitors.

For testing the hypotheses that the differentiation traits could be either (a) location, (b) quality, or (c) any other characteristics, an unsupervised random forest has been estimated in order to find out the variables of importance in setting the different segments or clusters of restaurants.

When analysing the variables in order of importance, the first 14 variables are related to schedule of the restaurant (open/closing hours); longitude and latitude -related to location- are in position 15th and 16th, with the review count being in position 17th. Quality, represented by the variable *stars* is 18th in the list of variable of importance to determine the clusters, followed by a list of several other attributes.

Then, we can assert that restaurants in Edinburgh are competing in a first round, in terms of schedules, while location is of secondary consideration, followed closely by quality, in third position.

Further analysis could then provide explanations wether these strategies are paying off in term of higher, or more frequent, ratings, or if clusters of competing restaurants are very disimilar between each others in terms of the rest of attributes. In a following stage, this project could try to evaluate the changes in the competing scene forecasting the opening or closing of new restaurants in a defined location.
